\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-15 manuscript/main.tex
twice differentiable basis

	\subsubsection{Finite-basis approximation of the predictors and regression coefficients}\label{sec: finite basis approximation}	
	Given a basis $\{b_l(t)\}$ of $L^2([0,1])$, the functional predictors and the corresponding regression coefficients can be decomposed into
	\begin{equation*}
		X_{ij}(t) = \sum_{l=1}^\infty \theta_{ijl} \, b_l(t), \quad 
		\beta_j(t) = \sum_{l=1}^\infty \eta_{0,jl} \, b_l(t),
	\end{equation*}
	where $\theta_{ijl}$ and $\eta_{0,jl}$ are the coefficients of $X_{ij}(t)$ and $\beta_j(t)$ corresponding to the $l$th basis function $b_l(t)$, respectively.
Since functional predictors are observed at finitely many discrete time points, we approximate them by truncating the expansion at $M$ terms, yielding a finite-dimensional projection. A moderately large $M$ (e.g., 15 or 20) suffices to capture rich functional variation, and no fine-tuning is needed, as our method controlls smoothness  via penalization (Section~\ref{sec: Regularization of PLS Components}). We denote the truncated space by
	\begin{equation}
		\mathcal{B} := \operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr).
	\end{equation}
	 We denote the  predictors projected onto $\mathcal{B}$ and the resulting coefficients as
	\begin{equation*}
		\widetilde{X}_{ij}(t) = \sum_{l=1}^M \theta_{ijl} \, b_l(t), \quad 
		\tilde{\beta}_j(t) = \sum_{l=1}^M \eta_{0,jl} \, b_l(t).
	\end{equation*}
	The $i$th  hybrid predictor with projected functional parts, is a tuple, denoted as
	\begin{equation*}
		\widetilde{W}_i := 
		(
		X_{i1}, \ldots, X_{iK}, \boldsymbol{Z}_i
		).
	\end{equation*}
	We adopt the following notation for the coefficients:
	\begin{equation}
		\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top \in \mathbb{R}^M,
		\quad
		\boldsymbol{\eta}_{0j} := (\eta_{0,j1}, \ldots, \eta_{0,jM})^\top  \in \mathbb{R}^M,
	\end{equation}
	and
	\begin{equation}
		\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M}.
	\end{equation}

The observations in the hybrid PLS problem can be compactly represented by the tuple
\begin{equation}\label{def:problem_instance}
	(B^\ast, \Theta, \mathbf{y}),
\end{equation}
defined as follows.
Let $B \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions, with entries  
\[
B_{m, m'} := \langle b_m, b_{m'} \rangle = \int_0^1 b_m(t)\, b_{m'}(t)\, dt,
\]
for $m, m' = 1, \ldots, M$.
Then define $B^\ast := \operatorname{blkdiag}(B, \ldots, B, I_p) \in \mathbb{R}^{(MK + p) \times (MK + p)}$,  
$\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}$,  
and $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
	\begin{remark}
While it is possible to use a different basis for each functional predictor, we adopt a common basis for all predictors to simplify the notation. However, the definition of $B^\ast$ remains general enough to accommodate different bases across the $K$ predictors.
\end{remark}

>>>>

	\subsubsection{Finite-basis approximation of the predictors and regression coefficients}\label{sec: finite basis approximation}	
	Given a twice-differentiable basis $\{b_l(t)\}$ of $L^2([0,1])$, the functional predictors and the corresponding regression coefficients can be decomposed into
	\begin{equation*}
		X_{ij}(t) = \sum_{l=1}^\infty \theta_{ijl} \, b_l(t), \quad 
		\beta_j(t) = \sum_{l=1}^\infty \eta_{0,jl} \, b_l(t),
	\end{equation*}
	where $\theta_{ijl}$ and $\eta_{0,jl}$ are the coefficients of $X_{ij}(t)$ and $\beta_j(t)$ corresponding to the $l$th basis function $b_l(t)$, respectively.
Since functional predictors are observed at finitely many discrete time points, we approximate them by truncating the expansion at $M$ terms, yielding a finite-dimensional projection. A moderately large $M$ (e.g., 15 or 20) suffices to capture rich functional variation, and no fine-tuning is needed, as our method controlls smoothness  via penalization (Section~\ref{sec: Regularization of PLS Components}). We denote the truncated space by
	\begin{equation}
		\mathcal{B} := \operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr).
	\end{equation}
	 We denote the  predictors projected onto $\mathcal{B}$ and the resulting coefficients as
	\begin{equation*}
		\widetilde{X}_{ij}(t) = \sum_{l=1}^M \theta_{ijl} \, b_l(t), \quad 
		\tilde{\beta}_j(t) = \sum_{l=1}^M \eta_{0,jl} \, b_l(t).
	\end{equation*}
	The $i$th  hybrid predictor with projected functional parts, is a tuple, denoted as
	\begin{equation*}
		\widetilde{W}_i := 
		(
		X_{i1}, \ldots, X_{iK}, \boldsymbol{Z}_i
		).
	\end{equation*}
	We adopt the following notation for the coefficients:
	\begin{equation}
		\boldsymbol{\theta}_{ij} := (\theta_{ij1}, \ldots, \theta_{ijM})^\top \in \mathbb{R}^M,
		\quad
		\boldsymbol{\eta}_{0j} := (\eta_{0,j1}, \ldots, \eta_{0,jM})^\top  \in \mathbb{R}^M,
	\end{equation}
	and
	\begin{equation}
		\Theta_j := (\boldsymbol{\theta}_{1j}, \ldots, \boldsymbol{\theta}_{nj})^\top  \in \mathbb{R}^{n \times M}.
	\end{equation}

The observations in the hybrid PLS problem can be compactly represented by the tuple
\begin{equation}\label{def:problem_instance}
	(B^\ast, \Theta, \mathbf{y}),
\end{equation}
defined as follows.
Let 
$B, B^{''} \in \mathbb{R}^{M \times M}$
denote the Gram matrices of the basis functions and their second derivatives, respectively, with entries  
\begin{equation}
\end{equation}
\[
B_{m, m'} := \langle b_m, b_{m'} \rangle = \int_0^1 b_m(t)\, b_{m'}(t)\, dt,
\]
for $m, m' = 1, \ldots, M$.
Then define $B^\ast := \operatorname{blkdiag}(B, \ldots, B, I_p) \in \mathbb{R}^{(MK + p) \times (MK + p)}$,  
$\Theta := (\Theta_1, \ldots, \Theta_K, Z) \in \mathbb{R}^{n \times (MK + p)}$,  
and $\mathbf{y} := (y_1, \ldots, y_n)^\top$.
	\begin{remark}
While it is possible to use a different basis for each functional predictor, we adopt a common basis for all predictors to simplify the notation. However, the definition of $B^\ast$ remains general enough to accommodate different bases across the $K$ predictors.
\end{remark}
